---
title: "ha2546_hw3"
author: "Hana Akbarnejad"
date: "4/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(viridis)
library(ggplot2)
library(readr)
library(patchwork)

library(caret)
library(glmnet)
library(MASS)
library(e1071)
library(pROC)
library(AppliedPredictiveModeling)
library(ISLR)



knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


\newpage

```{r data, include=FALSE}

data(Weekly)
weekly_data = Weekly
head(weekly_data)

weekly_data = weekly_data %>%
  janitor::clean_names() %>% 
  dplyr::select(direction, everything(), -today, - year)  # not including year (?)
```

### (a) Produce some graphical summaries of the Weekly data

Add some data description + binary outcome (direction)

```{r eda, echo=FALSE}

theme1 <- transparentTheme(trans = .4)
theme1$strip.background$col <- rgb(.0, .6, .2, .2)
trellis.par.set(theme1)
featurePlot(x = weekly_data[, 2:7],
y = weekly_data$direction,
scales = list(x=list(relation="free"),
y=list(relation="free")),
plot = "density", pch = "|",
auto.key = list(columns = 2))
```

Add some description of data

### (b) Use the full data set to perform a logistic regression with Direction as the response and the five Lag variables plus Volume as predictors. Do any of the predictors appear to be statistically significant? If so, which ones?

```{r logistic_reg}

```

